---
title: "Covid91 vaccine study Final2023"
author: "Andrea Calderon"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: Spring 2024 final 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


$H_o$: There is no difference in the effectiveness of the covid 921 vaccine in females.

$H_a$: There is a difference in the effectiveness of the covid 921 vaccine in females.

## Methods 

The variables in this study: Males-categorical, Females-categorical, LGBT-categorical, Druggies, categorical. We will be using the Cat~Cat helper file.



## Males



## Graphical Results 



## Females

```{r}
dataF = filter(FinalData,sex=="F")
barchartGC(~infected + treatment, data=dataF)
```

```{r}
barchartGC(~infected + treatment, data=dataF, type= "percent")
```

This graph shows us that mostly all the females that have contracted COVID, the females that received the vaccince expereinced less symptoms than those who have received the placebo. There is a big difference between those who have been vaccinated and not vaccinated. 


## Numerical Descriptive

```{r}
table1 <- xtabs(~infected+ treatment, data=dataF)
rowPerc(table1)

```


```{r}
colPerc(table1)
```


The information above represents that the vaccine has a lower infection rate compared to the placebo. The vaccination has a slightly higher rate of treatment success among individuals infected; the difference between the vaccine and placebo is.67. According to the information in this table, 60% of the female COVID-19 infected group who were not vaccinated showed symptoms, compared to 40% of the female group who received vaccinations. Furthermore, data indicates that 2.04% of the female placebo group and 1.37% of the vaccinated group respectively had COVID infections.


## Inferential Results 

```{r}
chisq.test(table1)
```


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The P-value is 0.0002732 for chi-squared test and there is a slight connection within the infected and treament for the female subgroup.

On the fisher exact test we have gotten the p-value of 0.0002162,it indicates the same of the chi-squared test. 

## LGBTQ

## Introduction 

$H_o$ : there is no difference in effectiveness of the covid vaccine between LGBTQ. 


$H_a$ : there is a difference in the effectiveness of the covid vaccine between LGBTQ

## Methods 

I will be using the cat_Cat rmd file as guidence

## Graphical Descriptive results 

```{r}
levels(FinalData$LGBTQ)
```

```{r}
dataG= filter(FinalData,LGBTQ=="gay")
barchartGC(~infected + treatment,data=dataG)
```


```{r}
barchart(~infected + treatment, data=dataG, type= "percent")
```

The vaccine and placebo differ significantly from one another, as this graphic demonstrates. Those who got the vaccination contracted COVID-19 at a higher rate than those who received a placebo. This graph demonstrates that among the LGBTQ subgroup, those who were infected with the COVID-19 virus and got vaccinations showed noticeably greater symptoms than those who did not. There was no difference observed between those who were vaccinated and those who were not, and who did not contract COVID.

## Numerical Results 

```{r}
table1 <- xtabs(~infected + treatment, data=dataG)
rowPerc(table1)
```

```{r}
colPerc(table1)
```

The LGBTQ placebo group experienced 15.38% of symptoms, however the vaccinated group experienced 84.62% of symptoms, as this table illustrates. Additionally, this graph demonstrates that whilst 3.54% of the vaccinated group contracted COVID, only 0.66% of the LGBTQ group receiving the placebo actually had COVID.


## Inferental Results 




## Druggies


# Overall Results and Conclusions