---
title: "Covid91 vaccine study Final2023"
author: "Andrea Calderon"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: Spring 2024 final 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


$H_o$: There is no difference in the effectiveness of the covid 921 vaccine in females.

$H_a$: There is a difference in the effectiveness of the covid 921 vaccine in females.

## Methods 

The variables in this study: Males-categorical, Females-categorical, LGBT-categorical, Druggies, categorical. We will be using the Cat~Cat helper file.



## Males



## Graphical Results 



## Females

```{r}
dataF = filter(FinalData,sex=="F")
barchartGC(~infected + treatment, data=dataF)
```

```{r}
barchartGC(~infected + treatment, data=dataF, type= "percent")
```

This graph shows us that mostly all the females that have contracted COVID, the females that received the vaccince expereinced less symptoms than those who have received the placebo. There is a big difference between those who have been vaccinated and not vaccinated. 


## Numerical Descriptive

```{r}
table1 <- xtabs(~infected+ treatment, data=dataF)
rowPerc(table1)

```


```{r}
colPerc(table1)
```


The information above represents that the vaccine has a lower infection rate compared to the placebo. The vaccination has a slightly higher rate of treatment success among individuals infected; the difference between the vaccine and placebo is.67. According to the information in this table, 60% of the female COVID-19 infected group who were not vaccinated showed symptoms, compared to 40% of the female group who received vaccinations. Furthermore, data indicates that 2.04% of the female placebo group and 1.37% of the vaccinated group respectively had COVID infections.


## Inferential Results 

```{r}
chisq.test(table1)
```


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The P-value is 0.0002732 for chi-squared test and there is a slight connection within the infected and treament for the female subgroup.

On the fisher exact test we have gotten the p-value of 0.0002162,it indicates the same of the chi-squared test. 

## LGBTQ

## Introduction 

$H_o$ : there is no difference in effectiveness of the covid vaccine between LGBTQ. 


$H_a$ : there is a difference in the effectiveness of the covid vaccine between LGBTQ

## Methods 

I will be using the cat_Cat rmd file as guidence

## Graphical Descriptive results 

```{r}
levels(FinalData$LGBTQ)
```

```{r}
dataG= filter(FinalData,LGBTQ=="gay")
barchartGC(~infected + treatment,data=dataG)
```


```{r}
barchart(~infected + treatment, data=dataG, type= "percent")
```

The vaccine and placebo differ significantly from one another, as this graphic demonstrates. Those who got the vaccination contracted COVID-19 at a higher rate than those who received a placebo. This graph demonstrates that among the LGBTQ subgroup, those who were infected with the COVID-19 virus and got vaccinations showed noticeably greater symptoms than those who did not. There was no difference observed between those who were vaccinated and those who were not, and who did not contract COVID.

## Numerical Results 

```{r}
table1 <- xtabs(~infected + treatment, data=dataG)
rowPerc(table1)
```

```{r}
colPerc(table1)
```

The LGBTQ placebo group experienced 15.38% of symptoms, however the vaccinated group experienced 84.62% of symptoms, as this table illustrates. Additionally, this graph demonstrates that whilst 3.54% of the vaccinated group contracted COVID, only 0.66% of the LGBTQ group receiving the placebo actually had COVID.


## Inferental Results 


```{r}
chisq.test(table1)
```


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


Since the p-value is zero, we can rule out the null hypothesis and come to the conclusion that there is a meaningful relationship between the type of treatment and the state of infection. The type of treatment is associated to contracting the virus, regardless of whether the recipient received the vaccine or a placebo. We are rejecting the null hypothesis based on the chisq test, which indicates that the p-value for the LGBTQ group is 0.0000006668. This indicates that there is an effect between the COVID vaccination and the LGBTQ subgroup.  

## Druggies

## Introduction 
$H_o$ : there is no difference in of COVID921 vaccine within druggies 

$H_a$ : there is no difference in the COVID921 vaccine within druggies

## Methods 

We will be using the Cat~cat helper file to guide me 


## Graphical Descriptive Results 

```{r}
levels(FinalData$Drugusers=="yes")
```

```{r}
dataD=filter(FinalData,DrugUser=="yes")
barchartGC(~infected +treatment, data=dataD)
```

```{r}
barchartGC(~infected + treatment, data=dataD, type= "percent")
```

This graph demonstrates that among the drug user group that contracted COVID, those who received the vaccine had more symptoms than those who received a placebo, but there was no difference between the vaccinated and unvaccinated groups in the drug user group that did not contract COVID.


## Numerical 

```{r}
table1 <- xtabs (~infected + treatment,data=dataD)
rowPerc(table1)
```


```{r}
colPerc(table1)
```


According to the data in this table, of the druggies group that contracted COVID-19, 11.67% of those who were not vaccinated and 88.33% of those who were vaccinated reported having symptoms. Additionally, it demonstrates that 3.06% of the vaccinated group developed COVID symptoms, and 0.4% of the druggies placebo group had COVID symptoms.


## Inferential Results 


```{r}
chisq.test(table1)
```

```{r}
chisqtestGC(table1)
```

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value on this chart is exactly the same as it was on the previous one. Similarly, because the p-value is 0, we will reject the null hypothesis. As a result, we shall draw the conclusion that treatment type and infection state are significantly correlated. We obtained a p-value of 0.000000000232 from the Fisher exact test, which is consistent with the results of the Chi-squared test.


# Overall Results and Conclusions


In result, there are considerable differences in the COVID vaccine's efficacy among various demographic groupings. The percentage of females who contracted COVID-19 was much lower in those who received the vaccine than in those who received a placebo. This finding was verified (p < 0.001) by the Fisher exact test and the Chi-squared test. In comparison to those who received a placebo, members of the LGBTQ subgroup exhibited a greater percentage of COVID-19 symptoms in those who received the vaccine. This unexpected outcome was validated by both statistical tests (p < 0.001). Addicts who received the vaccination had a greater risk of contracting COVID-19, much like the LGBTQ subgroup (p < 0.001).Thus, we reject the null hypothesis that the COVID vaccine is equally efficacious for individuals in different subgroups based on the statistical tests performed for each grouping.